Search for Spatially Extended Fermi-LAT Sources Using Two Years of Flight Data

### 2011-07-14 extended_catalog/paper/extended_catalog.tex
mroe chnages

Being able to spatially resolve the \gev emission from a source is
important for several reasons. Because of the large point spread
function (PSF), there are often several possible source counterparts.
Finding a matching morphology can be used to uniquely identify the source.
Because of the largely varying PSF of the LAT, the spatial and spectral
information about a source do not nicely decouple. Therefore, a biased
spatial model will bias the spectral model of the source and studyign
the extension of the source can correct for htis.  Furthermore, correctly
modeling a source's extension is important for improving the overall model
of the sky and removing excess residuals.  These residuals manifested
themseleves in the one year catalog as spurious point sources in the
region region surrounding the Large Magellanic Cloud (\cite{first_cat}).

@@

Errors on fit parameters are typically calculated in one of two
ways. The first method requires explicitly varying a parameter
while simultaneously optimizing the other parameters.
until the log of the likelihood has fallen by a particular value.
The \minuit fitting
package provides the {\em Simplex} function to calculate errors this way
(\cite{minuit_documentation}).  The second method involves estimating the
likelihood function as being a multivariate Gaussian and calculating the
curvature of the log of the likelihood at the peak to estimate when the
function will decrease by the desired value. \minuit provides the {\em HESSE}
algorithm to calculate errors this way.  {\em HESSE} errors 
can be calculated significantly faster.

@@

The spatial shape of extended soruces can e modeled
as a two dimensional Gaussian
\begin{equation}
  \text{PDF}(x,y)=\tfrac{1}{2\pi\sigma^2}\exp\left(-(x^2+y^2)/2\sigma^2\right)
\end{equation}
or a uniform disk
\begin{equation}
  \text{PDF}(x,y)=\tfrac{1}{\pi\sigma^2}\delta\left(x^2+y^2<\sigma^2\right).
\end{equation}
Although these shapes are significantly different, the convolution
of them with the PSF for typical sizes look very similar.  Even
though these shapes are significantly different, we demonstrate in
figure~\ref{compare_disk_gauss} that the convolution of shows the
PSF that would be observed for a power-law source of spectral index
2, the probability density function (PDF) for a Gaussian and uniform
disk of width $0.5\deg$, and the convolution of this PDF with the PSF.
The figure shows the convolution for the PSF in the energy range between
1 \gev and and 100 \gev and also in the energy range between 10 \gev and
100 \gev.  The convolved shapesdo separate better at higher energy, but
nevertheless this plot shows that we are not very sensitive to probing
the exact structure of an extended source.  In this paper we always
assume a uniform disk spatial model.

@@

The half $\chi^2_1$ distribution was found to hold for the special case
of source finding using EGRET (\cite{mattox_egret}). This distribution
is also plausable when testing point sources for extension. Due to
statistical flucuations, half of the point sources might look narrower
than the PSF leading to $\tsext=0$. The other half of the time, we would
get a distribution consistent with Wilk's theorem. We used a Monte Carlo
study to demonstrat that this is the observed distribution.
 We tested this hypothesis using a dedicated Monte
Carlo study.

The Monte Carlo study involved simulating point sources of varying
spectral models and fitting them using \pointlike as both point
and extended sources.  LAT data was simulated using program
\gtobssim\footnote{\gtobssim is distributed by and supported by the
Fermi Science Support Center (\cite{fssc}).}  The sources were simulated
with a power-law spectral model with 6 Were given 6 $>100$ \mev fluxes
ranging from $3\times 10^{-9} \ph/\cm^2/\sec$ to $10^{-6} \ph/\cm^2/\sec$.
The sources were then simulated with spectral index ($\gamma$) values of
1.5, 2, 2.5, and 3.  These values were picked to probe a representative
range measurable values.  For particular spectral indices, only the fluxes
which would create a statistically significant source detection ($\ts>25$)
where selected, since we are only interested in testing for extension
sources with a significant extent. The sources were simulated on top of
an isotropic power-law background with a Sreekumar-like spectrum ($>100$
\mev flux of 1.5e-5 and spectral index 2.1) (\cite{sreekumar_isotropic}).

The Monte Carlo simulation was performed over a one year time interval
using a default rocking profile and a representative livetime fraction
of 0.8.  The reconstruction was performed with photons from an energy
range of 1 \gev to 100 \gev using 4 energy bins per decade and using
the Pass 7 Version 6 source instrument response function (IRFs) to be
consistent with the following analysis.

Before we present the results of the Monte Carlo simulations, we have to
bring up one issue we found with these tests.  Despite our best efforts,
there was always a small numerical discrepancy between the likelihood when
fitting the source as the PSF and when fitting it as a very small extended
source with This is due to numerical error introduced by the convolution.
In most situations, this error would be insignificant. But it does bias
the distribution of \tsext in the null hypothesis. To correct for this,
instead of computing the likelihood in the null hypothesis by fitting
the PSF, we instead fit the null hypothesis as an extended source with
the extension fixed to be very small.

For most of the spectral models, over 10,000 statistically independent
simulations were performed.  For fainter sources, some of the
simulated sources, statistical fluctuations caused the source
to be not significant and they were excluded from the following
analysis.  Table~\ref{ts_ext_num_sims} shows the parameters used in
the simulation, the number of simulations performed, and the average
test statistic value.  The results of the Monte Carlo simulation are
shown in figure~\ref{ts_ext_mc}.  The cumulative density function (CDF)
is plotted against $\tsext$ For reference, the CDF for $\chi^2_1/2$,
suggested by Wilks' Theorem, is also plotted for comparison.

Our Monte Carlo study showed broad agreement between simulations and
Wilks' theorem and that we can use \tsext as a test of the statistical
significance of a source's extension.  Nevertheless, the agreement
is not perfect.  It should be noted that the discrepancy seems to be
worst for bright sources. This seems to imply that numerical errors in
the convolution become more apparent for sufficiently bright sources.
Other possible reasons for departure from Wilks' theorem might include
\pointlike ignoring energy dispersion. Energy dispersion will change
the shape as a function of energy. \pointlike may also run into possibly
problems with the fitting algorithm getting stuck in false minimum. It
is important to emphasize that if our empirical curve lies to the left of
the curve we assume to estimate significance, we would underestimate the
statistical significance of our result. Since this is the case for most
of our simulations, we are reasonable confident that $\sqrt{\tsext}$
can be safely used as a measure of the statistical significance of a
source's extension.

@@

We performed a second study to determine the LAT's detection threshold
to spatially extended sources. The detection threshold to extension
is defined as the flux at which the average value of
$\tsext$ is equal to $\langle\ts\rangle=25$. As was seen in
section~\ref{monte_carlo_validation}, this corresponds to an average
extension significance of $5\sigma$.
Qualitatively, we expect that the LAT would not be sensitive to
extended sources much smaller than the PSF. Similarly,
we expect that the LAT would not be sensitive to sources
much larger than the PSF since they would get lost in the
background.

To calculate the LAT's detection threshold to spatially extended sources,
we simulated sources with a uniform radially symmetric disk
morphology on top of a power-law Sreekumar-like isotropic spectrum
(\cite{sreekumar_isotropic}).  The Monte Carlo simulation was performed
over a one year time interval using a default rocking profile and a
livetime fraction of 0.8.  For a given extension and spectral index, we
first hand picked a flux range such that $\tsext<25$ and $\tsext>25$. We
then picked 10 flux points in this range and performed an extension test
for each of these fluxes. We repeated this procedure $>15$ times for each
spectral and spatial model.  By fitting a line to the flux vs $\tsext$,
we estimated when $\langle\tsext\rangle=25$.

Figure~\ref{index_sensitivity} shows the threshold for
sources of different spectral indices and extensions.  We selected
spectral indices 1.5, 2, 2.5, and 3 and 
extensions from $\sigma=0.1\deg$ to $2.0\deg$.  The curves
on the plot represent the division in extension and flux space where
$\langle\tsext=25\rangle$.  The plot shows the threshold using all
photons with energy between 100 \mev and 100 \gev and the threshold
using only photons with energy between 1 \gev and 100 \gev.  This results
shows that the LAT's threshold to extension is better for harder sources
with more high energy photons. The threshold is best for sources with
an extension around $0.5\deg$ and is worse for small and large sources.

Except for the hardest and largest sources, our detection threshold is almost
unchanged when using events between 100 \mev and 1 \gev.  This is also
demonstrated in figure~\ref{ts_ext_vs_energy} which shows the \tsext for
the SNR IC443 in 12 energy bins between 100 \mev and 100 \gev. For IC443,
which has a spectral index ~2.4, almost the entire increase in \ts comes
from energies above 1 \gev.  On the other hand, other systematic errors
become increasingly important at low energy. For our extension search,
we include only photons above 1 \gev.

Figure~\ref{diff_factor_sensitivity} shows how the LAT's
detection threshold varies with increasing background, typical of sources
at lower galactic latitude. We compute it for spectral
index 2 sources using 10 and 100 times the
extragalactic background. The latter result approximately represents the intensity of
the galactic diffuse emission near the galactic center. 
The figure shows two plots, the first showing the threshold looking only between 
1 \gev and 100 \gev and the second showing the threshold looking only
between 10 \gev and 100 \gev.
For each plot, the $y$ axis quotes the flux in the selected energy range
which weaken the dependence of the detection threshold on spectral index.
This shows that we are less sensitive to extended sources in
a higher background region and that if the flux is constant,
we are more sensitive if the source is harder (has more high energy
photons).

\section{Extension search method}

\todo[inline]{Add note about how we had to delete nearby sources
which were not hard above 10GeV for analysis to converge.}

@@

We performed a search of all Fermi sources found in the second Fermi LAT
catalog for extension.  To perform this search, we built a model of the
sky consistent with the two year catalog's model.   We used the same two
years of data from August 4, 2008 to August 1, 2010 and we used the same
pass 7 instrument response function.  We used the same galactic diffuse,
isotropic diffuse, and earth limb emission model. The spectrum of the
galactic and isotropic diffuse models were refit during the analysis
while the earth limb emission was fixed to it's predicted value.

@@

% 1FGL J1613.6-5100c - P72Y2473         - 2FGL J1615.2-5138
% Deleted
%  *P72Y2490 - 2FGL J1620.6-5111c
%  *P72Y2487 - 2FGL J1619.7-5040c
%  *P72Y2496 - 2FGL J1622.8-5006
% Modify 
%  * P72Y2472 - 2FGL J1615.0-5051 - SkyDir(332.410,-0.293,SkyDir.GALACTIC)

@@

 To get a good fit for the region, we had to remove from our
model the nearby catalog sources 2FGL J1520.6-5111c, 2FGL J1619.7-5040c,
and 2FGL J1622.8-5006. Furthermore, to get a good fit of the region we
iterativly relocalized the source 2FGL J1615.0-5051 while fitting the
position and extension of 2FGL J1615.2-5138.  We ended up with a best
fit position of the nearby source 2FGL J1615.0-5051 having moved from
$(l,b)=(332.2684,0.043)$ to $(l,b)=(332.41,-0.29)$. This was necessary
because the initial localization of this source was done assumign that
2FGL J1615.2-5138 was modeled as several point sources.  To address
the possibility of source confusion, to compare the overal change in
likelihood fitting 2FGL J1615.2-5138 as two point sources and find
\tsinc=34.3 which is smaller then \tsext.

@@

See figure~\ref{1FGL_J1613.6-5100c}.

@@

% 1FGL J1632.9-4802c - P72Y2540         - 2FGL J1632.4-4753c
% Deleted
% * P72Y2535 - 2FGL J1631.7-4720c
% * P72Y2550 - 2FGL J1636.3-4740c
% * P72Y2556 - 2FGL J1638.0-4703c
% * P72Y2528 - 2FGL J1630.2-4752
% * P72Y2543 - 2FGL J1634.4-4743c
% * P72Y2539 - 2FGL J1632.4-4820c
% * P72Y2521 - 2FGL J1628.1-4857c
% * P72Y2527 - 2FGL J1630.1-4615
% * P72Y2562 - 2FGL J1639.8-4921c
% Modify
% * P72Y2547 - 2FGL J1635.4-4717c - SkyDir(337.230,0.346,SkyDir.GALACTIC) 
%              (initial Position  - SkyDir(337.1396,0.1433)

@@

To get a good fit of the region, had to remove from our model the 5
catalog sources 2FGL J1631.7-4720c, 2FGL J1636.3-4740c, 2FGL J1630.2-4752,
2FGL J1634.4-4743c, and 2FGL J1632.4-4820c. These sources all are on top
of this extended source but there is no indication that the emission of
these sources separates.  These sources are indicated as green stars in
figure~\ref{1FGL_J1632.9-4802c}.

In addition we removed from our region the 4 nearby catalog sources
2FGL J1638.0-4703c, 2FGL J1628.1-4857c, 2FGL  J1630.1-4615, 2FGL
J1639.8-4921c because they were not significant above 10 \gev. Finally,
we had to move the source 2FGL J1635.4-4717c to a better fit position
$(l,b)=(337.230,0.346)$ to get a good fit of it and our extended source in
the $>10$ \gev energy range. The initial position of this source is shown
in the figure as a blue star and its best fit position is also indicated.

@@

Gamam Cygni SNR.

% 1FGL J2020.0+4049  - P72Y3281         - 2FGL J2021.5+4026
% Delete:
% * P72Y3282 - 2FGL J2019.1+4040
% * P72Y3287 - 2FGL J2020.0+4159
% * P72Y3262 - 2FGL J2013.8+4115c
% * P72Y3260 - 2FGL J2012.4+3955c
\todo[inline]{Describe how ROI was modified}

@@

% Delete:
% * P72Y2979 - 2FGL J1835.5-0649
% * P72Y2994 - 2FGL J1839.3-0558c
% * P72Y2993 - 2FGL J1839.0-0539

% Modify:
% * P72Y2985 - 2FGL J1836.8-0623c - SkyDir(25.574,0.320,SkyDir.GALACTIC)
%              (initial Position - SkyDir(,)
% * P72Y2974 - 2FGL J1837.3-0700c - SkyDir(24.716,0.500,SkyDir.GALACTIC)
%              (initial Position - SkyDir(,)

@@

% (no 1FGL) - P72Y1287         - 2FGL J0851.7-4635
% Delete:
% * P72Y1291 - 2FGL J0853.5-4711
% * P72Y1296 - 2FGL J0855.4-4625
% * P72Y1274 - 2FGL J0848.5-4535
% * P72Y1259 - 2FGL J0842.9-4721
% * P72Y1310 - 2FGL J0901.7-4655
% * P72Y1300 - 2FGL J0858.0-4815
% Modify:
%  * P72Y1293 - 2FGL J0854.7-4501 - SkyDir(266.232,0.517,SkyDir.GALACTIC))
%              (initial Position - SkyDir(265.5705,-0.0099)

@@

\begin{figure}
  \begin{center}
    \includegraphics{mc_plots/compare_disk_gauss.pdf}
    % this plot came from /u/gl/lande/work/fermi/extended_catalog/2FGL/plots_for_paper/compare_disk_gauss/v2/compare_disk_gauss.pdf
    \end{center}
    \caption{
    This plot demonstrates that different spatial models when smoothed
    by the PSF look very similar.  The $y$ axis is the log of the
    intensity and the $x$ axis is $\theta^2$.  The black line is the
    PSF that would be observed for a power-law source of spectral index
    2. The dashed red line is the probability density of a Gaussian with
    $\rsixeight=0.5\deg$, which is just a straight line on this plot
    The dashed blue line is the probability density for a uniform disk
    with $\rsixeight=0.5\deg$.  Finally, the solid red and blue lines
    show the convolution of the Gaussian and uniform disk with the PSF.
    The left plot shows the analysis with an energy range from 1 \gev to
    100 \gev whereas the right plot shows the analysis with an energy
    range from 10 \gev to 100 \gev.  This plot is discussed futher in
    section~\ref{compare_source_size}
    }\label{compare_disk_gauss}
  \end{figure}

  \begin{figure}
    \begin{center}
      \includegraphics{mc_plots/compare_r68.pdf}
      % this plot came from /u/gl/lande/work/fermi/extended_catalog/2FGL/plots_for_paper/compare_r68/v2/plot.py
      \end{center}
      \caption{
      This plot shows that fitting the same source with different spatial
      shapes can lead to very different values of $\sigma$. On the other
      hand, the different shapes have a good agreement about the 68\%
      containment radius \rsixeight.  It shows counts map for a simulated
      uniform disk extended source in the energy range from 1 \gev to
      100 \gev.  The source has a size $\sigma=0.5\deg$, an integral flux
      between 1 \gev and 100 \gev of $3\times 10^{-8}\ph/\cm^2/\sec$, and
      a spectral index 2.  The smoothed counts from a point source with
      the same spectrum are shown in the bottom right inset.  This source
      was fit with both a uniform disk and a Gaussian spatial model.
      The white circle represents the fit size $\sigma$ modeling the
      source with a uniform disk spatial model.  The black circle is
      the fit $\sigma$ modeling the source with a Gaussian hypothesis.
      The dashed white and black circles are the fit \rsixeight for the
      disk and Gaussian hypothesis. This plot is discussed furhter in
      section~\ref{compare_source_size}
      }\label{compare_r68}
    \end{figure}

>>>>

Being able to spatially resolve the \gev emission from a source is
important for several reasons. Because of the large point spread
function (PSF), there are often several possible source counterparts.
Finding a matching morphology can be used to uniquely identify the source.
Because of the largely varying PSF of the LAT, the spatial and spectral
information about a source do not nicely decouple. Therefore, a biased
spatial model will bias the spectral model of the source and studyign
the extension of the source can correct for htis.  Furthermore, correctly
modeling a source's extension is important for improving the overall model
of the sky and removing excess residuals.  These residuals manifested
themselves in the one year catalog as spurious point sources in the
region region surrounding the Large Magellanic Cloud (\cite{first_cat}).

@@

Errors on fit parameters are typically calculated in one of two
ways. The first method requires explicitly varying a parameter
while simultaneously optimizing the other parameters
until the log of the likelihood has fallen by a particular value.
The \minuit fitting
package provides the {\em Simplex} function to calculate errors this way
(\cite{minuit_documentation}).  The second method involves estimating the
likelihood function as being a multivariate Gaussian and calculating the
curvature of the log of the likelihood at the peak to estimate when the
function will decrease by the desired value. \minuit provides the {\em HESSE}
algorithm to calculate errors this way.  {\em HESSE} errors 
can be calculated significantly faster.

@@

The spatial shape of extended soruces can e modeled
as a two dimensional Gaussian
\begin{equation}
  \text{PDF}(x,y)=\tfrac{1}{2\pi\sigma^2}\exp\left(-(x^2+y^2)/2\sigma^2\right)
\end{equation}
or a uniform disk
\begin{equation}
  \text{PDF}(x,y)=\tfrac{1}{\pi\sigma^2}\delta\left(x^2+y^2<\sigma^2\right).
\end{equation}
Although these shapes are significantly different, the convolution
of them with the PSF for typical sizes look very similar.  Even
though these shapes are significantly different, we demonstrate in
figure~\ref{compare_disk_gauss} that the convolution of shows the
PSF that would be observed for a power-law source of spectral index
2, the probability density function (PDF) for a Gaussian and uniform
disk of width $0.5\deg$, and the convolution of this PDF with the PSF.
The figure shows the convolution for the PSF in the energy range between
1 \gev and 100 \gev and also in the energy range between 10 \gev and
100 \gev.  The convolved shapes do separate better at higher energy, but
nevertheless this plot shows that we are not very sensitive to probing
the exact structure of an extended source.  In this paper we always
assume a uniform disk spatial model.

@@

The half $\chi^2_1$ distribution was found to hold for the special case
of source finding using EGRET (\cite{mattox_egret}). This distribution
is also plausible when testing point sources for extension. Due to
statistical fluctuations, half of the point sources might look narrower
than the PSF leading to $\tsext=0$. The other half of the time, we would
get a distribution consistent with Wilk's theorem. We will present a
Monte Carlo study that shows an empirical distribution of \tsext in the
null hypothesis and compares it to this distribution.

We simulated point sources of varying spectral models and fit them
them using \pointlike as both point and extended sources. The data was
simulated using \gtobssim\footnote{The \gtobssim program is distributed
and supported by the Fermi Science Support Center (\cite{fssc}).}
The point sources were simulated with a power-law spectral model with
six $>100$ \mev fluxes ranging from $3\times 10^{-9} \ph/\cm^2/\sec$ to
$10^{-6} \ph/\cm^2/\sec$.  The point sources were simulated with spectral
indices ($\gamma$) of 1.5, 2, 2.5, and 3.  These values were picked to
to represent the types of sources detectable by the LAT.  The point
sources were simulated on top of an isotropic power-law background
with a Sreekumar-like spectrum ($>100$ \mev flux of 1.5e-5 and spectral
index 2.1) (\cite{sreekumar_isotropic}).  The Monte Carlo simulation was
performed over a one year time interval using a default rocking profile
and a representative livetime fraction of 0.8.  The reconstruction was
performed with photons from an energy range of 1 \gev to 100 \gev using 4
energy bins per decade and using the Pass 7 Version 6 Source instrument
response function (IRFs).  For each point source, we used \pointlike
to fit it as both a point and extended source and calculate \tsext and
we only dept the sources which had a significant point source detection
($\ts>25$).

Before we present the results of the Monte Carlo simulations, we note
that despite out best efforts, there was always a small numerical
discrepancy between the likelihood when fitting the source as the PSF
and when fitting it as a very small extended source. This is due to
numerical error introduced by the convolution.  In most situations, this
error would be insignificant. But when testing a source for extension,
this offset significantly bias the distribution of \tsext in the null
hypothesis. To correct for this, instead of computing the likelihood in
the null hypothesis by fitting the PSF, we instead calculated \tsext by
fitting the sources as an extended source and as an extended source with
its extension fixed to ${10^{-10}}\deg$.

For most of the spectral models, ~30,000 statistically independent
simulations were performed. For the dimmer spectral models, many of the
simulations had $\ts<25$ and were discarded.  Table~\ref{ts_ext_num_sims}
shows the different spectral models used in our study as well as the
number of simulations.  The cumulative density of \tsext is plotted in
figure~\ref{ts_ext_mc}. The $\chi^2_1/2$ distribution suggested by Wilks'
Theorem is overlaid for comparison.

Our Monte Carlo study shows broad agreement between simulations and Wilks'
theorem. Nevertheless, the agreement is not perfect.  It should be noted
that the discrepancy seems to be worst for bright sources which seems to
imply that numerical errors in the convolution become more apparent for
sufficiently bright sources.  Other possible reasons for departure from
Wilks' theorem might include \pointlike ignoring energy dispersion which
would change the PSF's shape as a function of energy. But we emphasize
that most of the time, our empirical curve lies to the left of the
theoretical curve so using the theoretical distribution would lead to an
underestimate of the statistical significance of a detection. Therefore,
we are confident that $\sqrt{\tsext}$ can be used as a measure of the
statistical significance of a source's extension and use it in the
following analysis.

@@

We performed a second study to determine the LAT's detection
threshold to spatially extended sources. The detection threshold
to extension is defined as the flux at which the average
value of $\tsext$ is $\langle\ts\rangle=25$. As was seen in
section~\ref{monte_carlo_validation}, this corresponds to an extension
significance of $5\sigma$.  Qualitatively, we expect the LAT would not
be sensitive to sources much smaller than the PSF or much larger than
the PSF.  Small sources would look too much like the PSF and larger
sources would get lost in the background.

To calculate the LAT's detection threshold to spatially extended
sources, we simulated sources with a uniform radially symmetric disk
spatial model on top of a power-law Sreekumar-like isotropic spectrum
(\cite{sreekumar_isotropic}).  The Monte Carlo simulation was performed
using the same one year time interval as in section~\ref{monte_carlo_validation}.
For each extension and spectral index, we
first picked a flux range which bracketed $\tsext=25$. We then picked
ten flux points in this range and repeatedly simulated and performed
extension tests of the extended sources.  For each flux, we performed
$>15$ independent simulations and extension fits.  We calculated
$\langle\tsext\rangle=25$ by fitting a line to all the flux and
$\tsext$ values and estimating the flux when $\tsext=25$.

Figure~\ref{index_sensitivity} shows the threshold for sources of four
spectral indices from 1.5 to 3 and extension varying from $\sigma=0.1\deg$
to $2.0\deg$.  The curves on the plot represent the flux for a given
extension where $\langle\tsext=25\rangle$.  The plot shows the threshold
using photons with energies between 100 \mev and 100 \gev and the
threshold using only photons with energies between 1 \gev and 100 \gev.
This results shows that the LAT's threshold to extension is better for
harder sources with more high energy photons. This result also shows
that the threshold is best for sources with an extension of $~0.5\deg$
and is significantly worse for very small extended small.

Except for the hardest and largest sources, our detection threshold
is almost unchanged when using events between 100 \mev and 1 \gev.
This is also demonstrated in figure~\ref{ts_ext_vs_energy} which shows
the \tsext for the SNR IC443 computed independently in 12 energy bins between 100 \mev and
100 \gev. For IC443, which has a spectral index ~2.4, almost the entire
increase in \ts comes from energies above 1 \gev.  On the other hand,
other systematic errors become increasingly important at low energy. For
our extension search, we use only only photons with energy above 1 \gev.

Figure~\ref{diff_factor_sensitivity} shows how the LAT's detection
threshold varies with increasing diffuse background. We compute the
sensitivity for a power-law source of spectral index 2 using 1, 10, and
100 times the extragalactic background. The latter factor approximately
represents the intensity of the galactic diffuse emission near the
galactic center.  The figure shows two plots, the first showing the
threshold using only photons with energies between 1 \gev and 100 \gev and
the second showing the threshold using only photons with energies between
10 \gev and 100 \gev. This plot demonstrates that we are less sensitive
to extended sources in a higher background region and that if the flux of
a source is constant (same number of photons), we would be more sensitive
to the extension of that source if those photons had a higher energy.

\section{Extended Source Search Method}

\todo[inline]{Add note about how we had to delete nearby sources which
were not hard above 10 \gev for analysis to converge.}

@@

We performed a search of all Fermi sources found in the second Fermi LAT
catalog for extension.  To perform this search, we built a model of the
sky consistent with the two year catalog's model.   We used the same two
years of data from August 4, 2008 to August 1, 2010 and we used the same
pass 7 Version 6 Source IRFs.  We used the same galactic diffuse,
isotropic diffuse, and earth limb emission model. The spectrum of the
galactic and isotropic diffuse models were refit during the analysis
while the earth limb emission was fixed to it's predicted value.

@@

% 1FGL J1613.6-5100c - P72Y2473         - 2FGL J1615.2-5138
% Deleted
%  *P72Y2490 - 2FGL J1620.6-5111c
%  *P72Y2487 - 2FGL J1619.7-5040c
%  *P72Y2496 - 2FGL J1622.8-5006
% Modify 
%  * P72Y2472 - 2FGL J1615.0-5051 - SkyDir(332.365,0.005,SkyDir.GALACTIC)
% New source
%  * SkyDir(332.426,-0.331,SkyDir.GALACTIC

@@

 To get a good fit for the region, we had to remove from our
model the nearby catalog sources 2FGL J1520.6-5111c, 2FGL J1619.7-5040c,
and 2FGL J1622.8-5006. Furthermore, to get a good fit of the region
we iterativly relocalized the source 2FGL J1615.0-5051 while fitting
the position and extension of 2FGL J1615.2-5138.  We ended up with
a best fit position of the nearby source 2FGL J1615.0-5051 having
moved from $(l,b)=(332.2684,0.043)$ to $(l,b)=(332.365,0.005)$.
In addition,we had to add a new source not in the catalog at the position
of $(l,b)=(332.426,-0.331)$. The moved and now source are shown in
figure~\ref{1FGL_J1613.6-5100c} are shown as the blue and XXXXXXXXXX\todo{what color} crosses
respectivly.  The location of these sources was found by iterativly
fitting the position of these sources as well as the extension of 2FGL
J1615.2-5138 until a stable fit was found.  It is unclear of this new seed
is physically a new source or if 2FGL J1615.0-5051 is spatially extended.
2FGL J1615.0-5051 was not found to be significantly extended using two
years of data but the fit of 2FGL J1615.2-5138 was clearly growing
too large to obsorb residual in the direction of 
J1615.2-5138 until we added the additional point source.

To address the possibility of source confusion, we compared the likelihood 
fitting 2FGL J1615.2-5138 as two point sources to one source.
We found that \tsinc=34.3 which is smaller then \tsext so we believe
that this is extended emission.

@@

@@

% 1FGL J1632.9-4802c - P72Y2540         - 2FGL J1632.4-4753c
% Kept:
% * P72Y2539 - 2FGL J1632.4-4820c
% Deleted
% * P72Y2535 - 2FGL J1631.7-4720c
% * P72Y2556 - 2FGL J1638.0-4703c
% * P72Y2528 - 2FGL J1630.2-4752
% * P72Y2543 - 2FGL J1634.4-4743c
% * P72Y2521 - 2FGL J1628.1-4857c
% * P72Y2527 - 2FGL J1630.1-4615
% * P72Y2562 - 2FGL J1639.8-4921c
% Modify
% * P72Y2547 - 2FGL J1635.4-4717c - SkyDir(337.230,0.346,SkyDir.GALACTIC)
%              (initial Position  - SkyDir(337.1396,0.1433)
% * P72Y2550 - 2FGL J1636.3-4740c - SkyDir(336.968,-0.066,SkyDir.GALACTIC))
%              (initial Position  - SkyDir(336.9639,-0.236)

@@

To get a good fit of the region, had to remove from our model the four
catalog sources 2FGL J1631.7-4720c, 2FGL J1630.2-4752,
2FGL J1634.4-4743c.4-4820c. These sources all are on top
of this extended source but there is no indication that the emission of
these sources separates.  These sources are indicated as green stars in
figure~\ref{1FGL_J1632.9-4802c}.

In addition we removed from our region the 4 nearby catalog sources 2FGL
J1638.0-4703c, 2FGL J1628.1-4857c, 2FGL  J1630.1-4615, 2FGL J1639.8-4921c
because they were not significant above 10 \gev.  Finally, we had to
relocalized the source 2FGL J1635.4-4717c to a better fit position
$(l,b)=(337.230,0.346)$ and the position of 2FGL J1636.3-4740c to the
better fit position $(l,b)=(336.968,-0.066)$ to get a good overall fit
of our source.

@@

Gamma Cygni SNR.

% 1FGL J2020.0+4049  - P72Y3281         - 2FGL J2021.5+4026
% Delete:
% * P72Y3292 - 2FGL J2022.8+3843c
% * P72Y3282 - 2FGL J2019.1+4040
% * P72Y3287 - 2FGL J2020.0+4159
% * P72Y3262 - 2FGL J2013.8+4115c
% * P72Y3260 - 2FGL J2012.4+3955c
% New source
%  * SkyDir(78.854,2.670,SkyDir.GALACTIC)

@@

% Delete:
% * P72Y2982 - Not in 2FGL
% * P72Y2979 - 2FGL J1835.5-0649
% * P72Y2993 - 2FGL J1839.0-0539

% Modify:
% * P72Y2985 - 2FGL J1836.8-0623c - SkyDir(25.574,0.320,SkyDir.GALACTIC)
%              (initial Position - SkyDir(25.5925,0.3103)
% * P72Y2974 - 2FGL J1837.3-0700c - SkyDir(24.716,0.500,SkyDir.GALACTIC)
%              (initial Position - SkyDir(25.0953,-0.0887)
% * P72Y2994 - 2FGL J1839.3-0558c - SkyDir(26.080,0.229,SkyDir.GALACTIC)
%              (initial Position - SkyDir(26.2301,-0.0394)

@@

% (no 1FGL) - P72Y1287         - 2FGL J0851.7-4635
% Delete:
% * P72Y1291 - 2FGL J0853.5-4711
% * P72Y1296 - 2FGL J0855.4-4625
% * P72Y1274 - 2FGL J0848.5-4535
% * P72Y1259 - 2FGL J0842.9-4721
% * P72Y1300 - 2FGL J0858.0-4815
% Modify:
% * P72Y1293 - 2FGL J0854.7-4501 - SkyDir(266.232,0.517,SkyDir.GALACTIC))
%              (initial Position - SkyDir(265.5705,-0.0099)
% * P72Y1310 - 2FGL J0901.7-4655 - SkyDir(267.311,-0.190,SkyDir.GALACTIC)
%              (initial Position - SkyDir(267.79,-0.3339)

@@

\begin{figure}
  \begin{center}
    \includegraphics{mc_plots/compare_disk_gauss.pdf}
    % this plot came from /u/gl/lande/work/fermi/extended_catalog/2FGL/plots_for_paper/compare_disk_gauss/v2/compare_disk_gauss.pdf
    \end{center}
    \caption{
    This plot demonstrates that different spatial models when smoothed
    by the PSF look very similar.  The $y$ axis is the log of the
    intensity and the $x$ axis is $\theta^2$.  The black line is the
    PSF that would be observed for a power-law source of spectral index
    2. The dashed red line is the probability density of a Gaussian with
    $\rsixeight=0.5\deg$, which is just a straight line on this plot
    The dashed blue line is the probability density for a uniform disk
    with $\rsixeight=0.5\deg$.  Finally, the solid red and blue lines
    show the convolution of the Gaussian and uniform disk with the PSF.
    The left plot shows the analysis with an energy range from 1 \gev to
    100 \gev whereas the right plot shows the analysis with an energy
    range from 10 \gev to 100 \gev.  This plot is discussed futher in
    section~\ref{compare_source_size}.
    }\label{compare_disk_gauss}
  \end{figure}

  \begin{figure}
    \begin{center}
      \includegraphics{mc_plots/compare_r68.pdf}
      % this plot came from /u/gl/lande/work/fermi/extended_catalog/2FGL/plots_for_paper/compare_r68/v2/plot.py
      \end{center}
      \caption{
      This plot shows that fitting the same source with different spatial
      shapes can lead to very different values of $\sigma$. On the other
      hand, the different shapes have a good agreement about the 68\%
      containment radius \rsixeight.  It shows counts map for a simulated
      uniform disk extended source in the energy range from 1 \gev to
      100 \gev.  The source has a size $\sigma=0.5\deg$, an integral flux
      between 1 \gev and 100 \gev of $3\times 10^{-8}\ph/\cm^2/\sec$, and
      a spectral index 2.  The smoothed counts from a point source with
      the same spectrum are shown in the bottom right inset.  This source
      was fit with both a uniform disk and a Gaussian spatial model.
      The white circle represents the fit size $\sigma$ modeling the
      source with a uniform disk spatial model.  The black circle is
      the fit $\sigma$ modeling the source with a Gaussian hypothesis.
      The dashed white and black circles are the fit \rsixeight for the
      disk and Gaussian hypothesis. This plot is discussed furhter in
      section~\ref{compare_source_size}.
      }\label{compare_r68}
    \end{figure}
